Privacy Analysis of a Mobile Sensor Application

### 2014-12-31 paper.tex
finished shortening of sensor privacy impact ouline

%\begin{LGContent}
In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
\textit{Privacy of the Person},
\textit{Privacy of Personal Behaviour},
\textit{Privacy of Personal Communication},
\textit{Privacy of Personal Data}.
It is important to note, that this categories do not form no taxonomy, since the categories are not mutually exclusive.
For instance a written email is considered personal communication as well as personal data stored on a computer.

@@

%\end{LGContent}

@@

\begin{figure}
\newcommand{\tcell}[1]{\parbox[c]{4cm}{\vspace{3mm}#1\vspace{3mm}}}
\newcommand{\thead}[1]{\tcell{\Large\centering\textbf{#1}}}
\newcommand{\tbody}[1]{\tcell{\Large\centering #1}}
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{|c||c|c|c|c|c|c|c|}
\hline
& \thead{Pricacy of the Person}
& \thead{Pricacy of Behaviour and Action}
& \thead{Pricacy of Communication}
& \thead{Pricacy of Data and Image}
& \thead{Pricacy of Thoughts and Feelings}
& \thead{Pricacy of Location and Space}
& \thead{Pricacy of Association}
\\\hline\hline
\thead{Geo Position}
& \tbody{0.5} % Pricacy of the Person
& \tbody{0.5} % Pricacy of Behaviour and Action
& \tbody{0} % Pricacy of Communication
& \tbody{1} % Pricacy of Data and Image
& \tbody{0} % Pricacy of Thoughts and Feelings
& \tbody{1} % Pricacy of Location and Space
& \tbody{0.5} % Pricacy of Association
\\\hline
\thead{Motion}
& \tbody{1} % Pricacy of the Person
& \tbody{0.5} % Pricacy of Behaviour and Action
& \tbody{0} % Pricacy of Communication
& \tbody{1} % Pricacy of Data and Image
& \tbody{0} % Pricacy of Thoughts and Feelings
& \tbody{0} % Pricacy of Location and Space
& \tbody{0} % Pricacy of Association
\\\hline
\thead{Networking}
& \tbody{0.5} % Pricacy of the Person
& \tbody{0.5} % Pricacy of Behaviour and Action
& \tbody{0} % Pricacy of Communication
& \tbody{1} % Pricacy of Data and Image
& \tbody{0} % Pricacy of Thoughts and Feelings
& \tbody{1} % Pricacy of Location and Space
& \tbody{1} % Pricacy of Association
\\\hline
%\thead{HAR}
%& \tbody{-} % Pricacy of the Person
%& \tbody{-} % Pricacy of Behaviour and Action
%& \tbody{-} % Pricacy of Communication
%& \tbody{-} % Pricacy of Data and Image
%& \tbody{-} % Pricacy of Thoughts and Feelings
%& \tbody{-} % Pricacy of Location and Space
%& \tbody{-} % Pricacy of Association
%\\\hline
%\thead{SLD}
%& \tbody{-} % Pricacy of the Person
%& \tbody{-} % Pricacy of Behaviour and Action
%& \tbody{-} % Pricacy of Communication
%& \tbody{-} % Pricacy of Data and Image
%& \tbody{-} % Pricacy of Thoughts and Feelings
%& \tbody{-} % Pricacy of Location and Space
%& \tbody{-} % Pricacy of Association
%\\\hline
\end{tabular}
}
\newline
\newline
{\scriptsize 
0: No Impact,
0.5: Light Impact,
1: Hard Impact}
\caption{Sensor Data Privacy Impact Matrix}
\label{figure:Sensor Data Privacy Impact Matrix}
\end{figure}

\subsubsection{GPS Data.}
The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use. Threefore, the collection of GPS data violates directly the citizens privacy of Location and Space, and the privacy of Data an Image.

\subsubsection{Motion Sensors.}
Accelerometer, Rotation Vector, Gyroscope and Magnetic field sensor measure the physical movement of the mobile device on all three axes.
If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
So his privacy is infringed regarding biometric behaviour, as it is captured automatically
Privacy of Data and Image is trivially threatened because here sensor data is individual data, a priori.

\subsubsection{Network Sensors.}
The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases.
The GSM sensor gives the exact cell, the mobile device has registered with at the current moment.

@@

>>>>

In Fried's definition of privacy as control over information, the specification of what constitues such information remains open.
There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
\textit{Privacy of the Person},
\textit{Privacy of Personal Behaviour},
\textit{Privacy of Personal Communication},
\textit{Privacy of Personal Data}.
It is important to note, that these categories do not form a taxonomy, since the categories are not mutually exclusive.
For instance a written email is considered personal communication as well as personal data stored on a computer.

Moreover, Friedewald et al. argue that Clarke's types are outdated and no longer adequate in order to describe the privacy aspect of our modern, technology driven, world.
In order to address this shortcoming they extend the former four to the now introduced seven types privacy as follows:

\begin{LGContent-Hidden}
In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
\textit{Privacy of the Person},
\textit{Privacy of Personal Behaviour},
\textit{Privacy of Personal Communication},
\textit{Privacy of Personal Data}.
It is important to note, that this categories do not form no taxonomy, since the categories are not mutually exclusive.
For instance a written email is considered personal communication as well as personal data stored on a computer.

@@

@@

Modern mobile devices have a broad collection of sensors.
Disclosure or processing of sensor data can impact one's privacy.
In this section we identify groups of sensors and their potential impact on a certain type of privacy.
Figure \ref{figure:Sensor Data Privacy Impact Matrix} qualifies that impact on a simple scale.
Privacy of Data and Image is trivially threatened because here sensor data is individual data, a priori.
Indirect impact is caused by combining sensor data with additional knowledge.
For instance, comparing a contemporary map with locational data can imply behaviour if the position matches a church.

\begin{figure}
\newcommand{\tcell}[1]{\parbox[c][2cm]{4cm}{\vspace{3mm}#1\vspace{3mm}}}
\newcommand{\thead}[1]{\tcell{\Large\centering\textbf{#1}}}
\newcommand{\tbody}[1]{\tcell{\Large\centering #1}}
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{|c||c|c|c|c|c|c|c|}
\hline
& \thead{Pricacy of the Person}
& \thead{Pricacy of Behaviour and Action}
& \thead{Pricacy of Communication}
& \thead{Pricacy of Data and Image}
& \thead{Pricacy of Thoughts and Feelings}
& \thead{Pricacy of Location and Space}
& \thead{Pricacy of Association}
\\\hline\hline
\thead{GPS Sensor}
& \tbody{0.5} % Pricacy of the Person
& \tbody{0.5} % Pricacy of Behaviour and Action
& \tbody{0} % Pricacy of Communication
& \tbody{1} % Pricacy of Data and Image
& \tbody{0} % Pricacy of Thoughts and Feelings
& \tbody{1} % Pricacy of Location and Space
& \tbody{0.5} % Pricacy of Association
\\\hline
\thead{Motion Sensors}
& \tbody{1} % Pricacy of the Person
& \tbody{0.5} % Pricacy of Behaviour and Action
& \tbody{0} % Pricacy of Communication
& \tbody{1} % Pricacy of Data and Image
& \tbody{0} % Pricacy of Thoughts and Feelings
& \tbody{0} % Pricacy of Location and Space
& \tbody{0} % Pricacy of Association
\\\hline
\thead{Networking Sensors}
& \tbody{0.5} % Pricacy of the Person
& \tbody{0.5} % Pricacy of Behaviour and Action
& \tbody{0} % Pricacy of Communication
& \tbody{1} % Pricacy of Data and Image
& \tbody{0} % Pricacy of Thoughts and Feelings
& \tbody{1} % Pricacy of Location and Space
& \tbody{1} % Pricacy of Association
\\\hline
%\thead{HAR}
%& \tbody{-} % Pricacy of the Person
%& \tbody{-} % Pricacy of Behaviour and Action
%& \tbody{-} % Pricacy of Communication
%& \tbody{-} % Pricacy of Data and Image
%& \tbody{-} % Pricacy of Thoughts and Feelings
%& \tbody{-} % Pricacy of Location and Space
%& \tbody{-} % Pricacy of Association
%\\\hline
%\thead{SLD}
%& \tbody{-} % Pricacy of the Person
%& \tbody{-} % Pricacy of Behaviour and Action
%& \tbody{-} % Pricacy of Communication
%& \tbody{-} % Pricacy of Data and Image
%& \tbody{-} % Pricacy of Thoughts and Feelings
%& \tbody{-} % Pricacy of Location and Space
%& \tbody{-} % Pricacy of Association
%\\\hline
\end{tabular}
}
\newline
\newline
{\scriptsize 
0: No Impact,
0.5: Indirect Impact,
1: Direct Impact}
\caption{Sensor Data Privacy Impact Matrix}
\label{figure:Sensor Data Privacy Impact Matrix}
\end{figure}

\subsubsection{GPS Sensor.}
The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use. Therefore, the collection of GPS data violates directly the citizens privacy of Location and Space.

%By inference we also get implicit violations of the Privacy of the Person, Privacy of Behavior and Action and Privacy of Association.

\subsubsection{Motion Sensors.}
Accelerometer, Rotation Vector, Gyroscope and Magnetic field sensor measure the physical movement of the mobile device on all three axes.
If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
So his privacy is infringed regarding biometric behaviour, i.e. the Privacy of the Person.

%By inference we also get implicit violations of the Privacy of Behavior and Action.

\subsubsection{Network Sensors.}
The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases.
The both sensors give the exact cell or network, the mobile device has registered with at the current moment.
Frequent connection to one particular network also reveals association, e.g. university networks.

@@

%A similar argumentat applies to WLAN snsors. 
%If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
